---
title: "AGL: Exp1"  
date: "`r Sys.Date()`"  
output:  
  word_document:  
    toc: true  
  html_document:  
    toc: true  
    theme: cosmo  
    df_print: paged  
---
  
```{r setup, include = FALSE}
knitr::opts_chunk$set(comment = "", error = TRUE)
```

# Load relevant libraries 

```{r, message = FALSE}
library(readxl)
library(stargazer)
library(car)
library(reshape)
library(plyr)
library(ggplot2)
library(afex)
library(dplyr)
library(tidyr)
library(psych)
```


# Set working directory for PC

To check the current working directory path 

```{r}
getwd()
```

# Load the data

```{r}
dataPs <- read_excel("AGL_Exp2.xlsx", sheet = "Behavioural")
dataStimuli <- read_excel("AGL_Exp2.xlsx", sheet = "ChunkStrength")

dataPassive <- read_excel("AGL_Exp2.xlsx", sheet = "PassiveExp2_predictions")
dataActive <- read_excel("AGL_Exp2.xlsx", sheet = "ActiveExp2_predictions")

dataIndPassive <- read_excel("AGL_Exp2.xlsx", sheet = "IndPassive")
dataIndActive <- read_excel("AGL_Exp2.xlsx", sheet = "IndActive")
```

## variables

age
:   Age in years

gender
:   Gender coded as male = 1, female = 2

group
:   Group coded as control group = 1, experimental group = 2

write
:   The percentage of correct responses in the writing task (only applicable to the experimental group)

gH
:   The endorsement rates for grammatical high-chunk items 

gL
:   The endorsement rates for grammatical low-chunk items

nH
:   The endorsement rates for non-grammatical high-chunk items

nL
:   The endorsement rates for non-grammatical low-chunk items

id
:   ID Number of Participants 

testItems
:   The letter strings presented in the testing phase 

chunkStrength
:   The chunk strength of each test item

category
:   The test items were categorized into four groups: grammatical and high chunk strength (1), grammatical and low chunk strength (2), non-grammatical and high chunk strength (3), and non-grammatical and low chunk strength (4)

```{r}
names(dataPs)
dim(dataPs)
View(dataPs)
```

```{r}
names(dataStimuli)
dim(dataStimuli)
View(dataStimuli)
```

### Age

```{r}
table(dataPs$age)
class(dataPs$age)
max(dataPs$age) 
min(dataPs$age)  
mean(dataPs$age)  
sd(dataPs$age)     
``` 

### Gender

```{r}
table(dataPs$gender)
class(dataPs$gender)
dataPs$gender.fa <- factor(dataPs$gender, levels=c(1:2), labels=c("Male","Female"))
table(dataPs$gender.fa)
class(dataPs$gender.fa)
```

### Group

```{r}
table(dataPs$group)
class(dataPs$group)
dataPs$group.fa <- factor(dataPs$group, levels=c(1:2), labels=c("Passive","Active"))
table(dataPs$group.fa)
class(dataPs$group.fa)
```

### Writing down task

```{r}
table(dataPs$write)
class(dataPs$write)
mean(dataPs$write, na.rm=TRUE)
```

### Grammatical high-chunk items

```{r}
table(dataPs$gH)
class(dataPs$gH)
table(dataPs$gH,dataPs$group.fa)
```

### Grammatical low-chunk items

```{r}
table(dataPs$gL)
class(dataPs$gL)
table(dataPs$gL,dataPs$group.fa)
```

### Non-grammatical high-chunk items

```{r}
table(dataPs$ngH)
class(dataPs$ngH)
table(dataPs$ngH,dataPs$group.fa)
```

### Non-grammatical low-chunk items

```{r}
table(dataPs$ngL)
class(dataPs$ngL)
table(dataPs$ngL,dataPs$group.fa)
```

### test items

```{r}
table(dataStimuli$testItems)
class(dataStimuli$testItems)
```

### Chunk Strength

```{r}
table(dataStimuli$chunkStrength)
class(dataStimuli$chunkStrength)
```

### Category

```{r}
table(dataStimuli$category)
class(dataStimuli$category)
dataStimuli$RuleAdherence_ChunkStrength <- factor(dataStimuli$category, levels=c("gH", "gL", "ngH", "ngL"), labels=c("g_H", "g_L", "ng_H", "ng_L"))
table(dataStimuli$RuleAdherence_ChunkStrength)
class(dataStimuli$RuleAdherence_ChunkStrength)
```

# Test Items and Chunk Strength in Experiment

```{r}
dataStimuli <- separate(dataStimuli, RuleAdherence_ChunkStrength, into = c("RuleAdherence", "ChunkStrength"), sep = "_")
```

#### Pair 1: compare GH and NGH

```{r}
gH <- subset(dataStimuli, RuleAdherence== "g" & ChunkStrength == "H")$chunkStrength
ngH <- subset(dataStimuli, RuleAdherence== "ng" & ChunkStrength == "H")$chunkStrength
```

```{r}
high_df <- data.frame(chunkStrength = c(gH, ngH), Group = c(rep("gH", length(gH)), rep("ngH", length(ngH))))
```

```{r}
t.test(chunkStrength ~ Group, data =high_df) # Welch by default
```

#### Pair 2: compare GL and NGL

```{r}
gL <- subset(dataStimuli, RuleAdherence== "g" & ChunkStrength == "L")$chunkStrength
ngL <- subset(dataStimuli, RuleAdherence== "ng" & ChunkStrength == "L")$chunkStrength
```

```{r}
low_df <- data.frame(chunkStrength = c(gL, ngL),Group = c(rep("gL", length(gL)), rep("ngL", length(ngL))))
```

```{r}
t.test(chunkStrength ~ Group, data =low_df) 
```

## compare high to low chunk strength within G and within NG
#### Pair 1: Within G

```{r}
gL <- subset(dataStimuli, RuleAdherence == "g" & ChunkStrength == "L")$chunkStrength
gH <- subset(dataStimuli, RuleAdherence == "g" & ChunkStrength == "H")$chunkStrength

t.test(gL, gH)  
```

#### Pair 2: Within NG

```{r}
ngL <- subset(dataStimuli, RuleAdherence == "ng" & ChunkStrength == "L")$chunkStrength
ngH <- subset(dataStimuli, RuleAdherence == "ng" & ChunkStrength == "H")$chunkStrength

t.test(ngL, ngH)
```

# Mixed-design ANOVA

```{r}
dataPs <- rename(dataPs, c("g_H" = "gH","g_L" = "gL","ng_H" = "ngH", "ng_L" = "ngL"))
names(dataPs)
```

```{r}
df_long <- pivot_longer(data = dataPs, cols = c(g_H, g_L, ng_H, ng_L), names_to = "RuleAdherence_ChunkStrength", values_to = "Endorsement")
```

```{r}
df_long <- separate(df_long, RuleAdherence_ChunkStrength, into = c("RuleAdherence", "ChunkStrength"), sep = "_")
```

```{r}
rule_adherence_levels <- c("g" = "Grammatical", "ng" = "Nongrammatical")
chunk_strength_levels <- c("H" = "High", "L" = "Low")
df_long$RuleAdherence <- factor(df_long$RuleAdherence, levels = names(rule_adherence_levels), labels = rule_adherence_levels)
df_long$ChunkStrength <- factor(df_long$ChunkStrength, levels = names(chunk_strength_levels), labels = chunk_strength_levels)
```

Check for NA values in the data

```{r}
anyNA(df_long)
View(df_long)
```

Check the number of levels in each factor

```{r}
print(paste("Number of levels in RuleAdherence:", length(levels(df_long$RuleAdherence))))
print(paste("Number of levels in ChunkStrength:", length(levels(df_long$ChunkStrength))))
print(paste("Number of levels in group:", length(levels(df_long$group.fa))))
```

## Performing factorial repeated measures ANOVA

```{r}
result <- aov_car(Endorsement ~ group.fa * RuleAdherence * ChunkStrength + Error(id/(RuleAdherence*ChunkStrength)), data = df_long)
summary(result)
```

## Post-hoc tests
### Pairwise t-tests

```{r}
pairwise.t.test(df_long$Endorsement, interaction(df_long$group.fa, df_long$RuleAdherence), p.adjust.method = "bonferroni")
```

## Calculate means, standard errors and standard deviation

```{r}
df_summary <- df_long %>%
  group_by(RuleAdherence, ChunkStrength, group.fa) %>%
  summarise(mean = mean(Endorsement, na.rm = TRUE),
            sd = sd(Endorsement, na.rm = TRUE),
            se = sd(Endorsement, na.rm = TRUE) / sqrt(n()), 
            .groups = "drop")
print(df_summary)
```

## Plot 1

```{r}
Violin_exp1 <- ggplot(df_long,
       aes(x = RuleAdherence,
           y = Endorsement * 100,
           fill = ChunkStrength)) +

  geom_violin(
    alpha = 0.45,
    colour = NA,
    width = 0.75,
    position = position_dodge(width = 0.7),
    trim = FALSE
  ) +

  geom_point(
    aes(color = ChunkStrength),
    position = position_jitterdodge(
      jitter.width = 0.10,
      jitter.height = 0,
      dodge.width = 0.7
    ),
    size = 1.4,
    alpha = 0.45
  ) +

  geom_boxplot(
    width = 0.12,
    outlier.shape = NA,
    colour = "black",
    fill = "white",
    linewidth = 0.35,
    position = position_dodge(width = 0.7)
  ) +

  stat_summary(
    fun = mean,
    geom = "point",
    position = position_dodge(width = 0.7),
    size = 2.6,
    colour = "black"
  ) +

  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    width = 0.12,
    colour = "black",
    linewidth = 0.6,
    position = position_dodge(width = 0.7)
  ) +

  facet_wrap(
    ~ group.fa,
    labeller = as_labeller(c(
      "Active" = "Active Learning Condition",
      "Passive" = "Passive Learning Condition"
    ))
  ) +

  labs(
    title = "(a) Experiment 1",
    x = "Grammaticality Status",
    y = "Endorsement rate (%)",
    fill = "Chunk strength",
    color = "Chunk strength"
  ) +

  scale_fill_manual(values = c("#4C78A8", "#F58518")) +
  scale_color_manual(values = c("#4C78A8", "#F58518")) +

  scale_y_continuous(
    limits = c(0,100),
    breaks = seq(0,100,20),
    expand = expansion(mult = c(0,0.02))
  ) +

  theme_minimal(base_size = 14) +
  theme(
    # Text
    plot.title = element_text(size = 14, face = "bold"),
    strip.text = element_text(size = 12),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 12),

    axis.line = element_blank(),

    strip.background = element_blank(),
    
    panel.border = element_blank(),

    panel.grid.major.y = element_line(colour = "grey90", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),

    legend.title = element_text(size = 13),
    legend.text = element_text(size = 12),
    legend.position = "right"
  )

print(Violin_exp1)

ggsave("Violin_exp1.tiff", Violin_exp1, width = 8, height = 5, dpi = 600, compression = "lzw")
```

# Endorsement rates 
## All 16 grammatical items
### Passive learners Group

```{r}
control_g <- subset(df_long, group.fa == "Passive" & RuleAdherence == "Grammatical")$Endorsement

mean_control_g <- mean(control_g, na.rm = TRUE)
print(paste("Endorsement Mean:", round(mean_control_g,3)))

sd_control_g <- sd(control_g, na.rm = TRUE)
print(paste("SD:", round(sd_control_g ,3)))

sem_control_g <- sd_control_g  / sqrt(length(na.omit(control_g)))
print(paste("SEM:", round(sem_control_g, 3)))
```

### Active learners Group

```{r}
experimental_g <- subset(df_long, group.fa == "Active" & RuleAdherence == "Grammatical")$Endorsement

mean_experimental_g <- mean(experimental_g, na.rm = TRUE)
print(paste("Endorsement Mean:", round(mean_experimental_g, 3)))

sd_experimental_g <- sd(experimental_g, na.rm = TRUE)
print(paste("SD:", round(sd_experimental_g, 3)))

sem_experimental_g <- sd_experimental_g / sqrt(length(na.omit(experimental_g)))
print(paste("SEM:", round(sem_experimental_g, 3)))
```

## All 16 nongrammatical items
### Passive learners Group

```{r}
control_ng <- subset(df_long, group.fa == "Passive" & RuleAdherence == "Nongrammatical")$Endorsement

mean_control_ng <- mean(control_ng, na.rm = TRUE)
print(paste("Endorsement Mean:", round(mean_control_ng, 3)))

sd_control_ng <- sd(control_ng, na.rm = TRUE)
print(paste("SD:", round(sd_control_ng, 3)))

sem_control_ng <- sd_control_ng / sqrt(length(na.omit(control_ng)))
print(paste("SEM:", round(sem_control_ng, 3)))
```

### Active learners Group

```{r}
experimental_ng <- subset(df_long, group.fa == "Active" & RuleAdherence == "Nongrammatical")$Endorsement

mean_experimental_ng <- mean(experimental_ng, na.rm = TRUE)
print(paste("Endorsement Mean:", round(mean_experimental_ng, 3)))

sd_experimental_ng <- sd(experimental_ng, na.rm = TRUE) 
print(paste("SD:", round(sd_experimental_ng, 3)))

sem_experimental_ng <- sd_experimental_ng / sqrt(length(na.omit(experimental_ng)))
print(paste("SEM:", round(sem_experimental_ng, 3)))
```

## All 16 high chunk-strength items
### Passive learners Group

```{r}
control_h <- subset(df_long, group.fa == "Passive" & ChunkStrength == "High")$Endorsement

mean_control_h <- mean(control_h, na.rm = TRUE)
print(paste("Endorsement Mean:", round(mean_control_h, 3)))

sd_control_h <- sd(control_h, na.rm = TRUE)
print(paste("SD:", round(sd_control_h, 3)))

sem_control_h <- sd_control_h / sqrt(length(na.omit(control_h)))
print(paste("SEM:", round(sem_control_h, 3)))
```

### Active learners Group

```{r}
experimental_h <- subset(df_long, group.fa == "Active" & ChunkStrength == "High")$Endorsement

mean_experimental_h <- mean(experimental_h, na.rm = TRUE)
print(paste("Endorsement Mean:", round(mean_experimental_h, 3)))

sd_experimental_h <- sd(experimental_h, na.rm = TRUE) 
print(paste("SD:", round(sd_experimental_h, 3)))

sem_experimental_h <- sd_experimental_h / sqrt(length(na.omit(experimental_h)))
print(paste("SEM:", round(sem_experimental_h, 3)))
```

## All 16 low chunk-strength items
### Passive learners Group

```{r}
control_l <- subset(df_long, group.fa == "Passive" & ChunkStrength == "Low")$Endorsement

mean_control_l <- mean(control_l, na.rm = TRUE)
print(paste("Endorsement Mean:", round(mean_control_l, 3)))

sd_control_l <-  sd(control_l, na.rm = TRUE)
print(paste("SD:", round(sd_control_l, 3)))

sem_control_l <- sd_control_l / sqrt(length(na.omit(control_l)))
print(paste("SEM:", round(sem_control_l, 3)))
```

### Active learners Group

```{r}
experimental_l <- subset(df_long, group.fa == "Active" & ChunkStrength == "Low")$Endorsement

mean_experimental_l <- mean(experimental_l, na.rm = TRUE)
print(paste("Endorsement Mean:", round(mean_experimental_l, 3)))

sd_experimental_l <- sd(experimental_l, na.rm = TRUE)
print(paste("SD:", round(sd_experimental_l, 3)))

sem_experimental_l <- sd_experimental_l / sqrt(length(na.omit(experimental_l)))
print(paste("SEM:", round(sem_experimental_l, 3)))
```

# Single-sample t-tests 
## Calculate grammaticality based on RuleAdherence and Endorsement

```{r}
df_long$grammaticality <- ifelse(df_long$RuleAdherence == "Grammatical", df_long$Endorsement, 1 - df_long$Endorsement)
View(df_long)
```

```{r}
average_grammaticality <- df_long %>%
  group_by(id) %>%
  summarise(average_grammaticality = mean(grammaticality, na.rm = TRUE))
average_grammaticality
```

## Create a new dataframe with group, id, and average grammaticality

```{r}
new_df <- df_long %>%
  select(group.fa, id) %>%
  distinct() %>%
  left_join(average_grammaticality, by = "id")

View(new_df)
```

## Calculate means, standard errors and standard deviation

```{r}
df_summary_grammaticality <- new_df %>%
  group_by(group.fa) %>%
  summarise(mean = mean(average_grammaticality, na.rm = TRUE),
            sd = sd(average_grammaticality, na.rm = TRUE),
            se = sd(average_grammaticality, na.rm = TRUE) / sqrt(n()), 
            .groups = "drop")
print(df_summary_grammaticality)
```

## Runing one Sample t-test to investigate mean classification accuracy for each group from chance level 

```{r}
t_test_passive <- t.test(new_df$average_grammaticality[new_df$group.fa == "Passive"], mu = 0.50, alternative = "greater")

t_test_active <- t.test(new_df$average_grammaticality[new_df$group.fa == "Active"], mu = 0.50, alternative = "greater")

t_test_passive
t_test_active
```

## Comparative Analysis of Classification Accuracy: Passive vs. Active Learning Conditions

```{r}
t.test(average_grammaticality ~ group.fa, data = new_df)
```

# RSS
## Min–max normalise: chunkStrength

```{r}
# Passive data
dataPassive$chunkStrength_norm <- (
  dataPassive$chunkStrength - min(dataPassive$chunkStrength)
) / (max(dataPassive$chunkStrength) - min(dataPassive$chunkStrength))

# Active data
dataActive$chunkStrength_norm <- (
  dataActive$chunkStrength - min(dataActive$chunkStrength)
) / (max(dataActive$chunkStrength) - min(dataActive$chunkStrength))
```

## Define an RSS function

```{r}
rss <- function(observed, predicted) {
  sum((observed - predicted)^2)
}
```

### Compute RSS for passive learning 

```{r}
RSS_Bayesian <- rss(dataPassive$Observed, dataPassive$Bayesian_predicted)
RSS_Quantum <- rss(dataPassive$Observed, dataPassive$Quantum_predicted)
RSS_chunkStrength <- rss(dataPassive$Observed, dataPassive$chunkStrength_norm)
RSS_grammaticality <- rss(dataPassive$Observed, dataPassive$grammaticality)
```

```{r}
RSS_results_Passive <- data.frame(
  Predictor = c("Bayesian", "Quantum", "Chunk Strength (norm)", "Grammaticality"),
  RSS = c(RSS_Bayesian, RSS_Quantum, RSS_chunkStrength, RSS_grammaticality)
)

RSS_results_Passive
```

### Compute RSS for Active learning 

```{r}
RSS_Bayesian_A <- rss(dataActive$Observed, dataActive$Bayesian_predicted)
RSS_Quantum_A <- rss(dataActive$Observed, dataActive$Quantum_predicted)
RSS_chunkStrength_A <- rss(dataActive$Observed, dataActive$chunkStrength_norm)
RSS_grammaticality_A <- rss(dataActive$Observed, dataActive$grammaticality)
```

```{r}
RSS_results_Active <- data.frame(
  Predictor = c("Bayesian", "Quantum", "Chunk Strength (norm)", "Grammaticality"),
  RSS = c(RSS_Bayesian_A, RSS_Quantum_A, RSS_chunkStrength_A, RSS_grammaticality_A)
)

RSS_results_Active
```

# Plot 2

### Create the category variable

```{r}
# Passive
dataPassive$category <- cut(
  seq_len(nrow(dataPassive)),
  breaks = c(0, 8, 16, 24, 32),
  labels = c("GH", "GL", "NGH", "NGL")
)

# Active
dataActive$category <- cut(
  seq_len(nrow(dataPassive)),
  breaks = c(0, 8, 16, 24, 32),
  labels = c("GH", "GL", "NGH", "NGL")
)
```


```{r}
# Passive
dataPassive <- dataPassive %>%
  group_by(category) %>%
  arrange(chunkStrength, .by_group = TRUE) %>%
  mutate(item_ordered = factor(testItem, levels = testItem))

# Active
dataActive <- dataActive %>%
  group_by(category) %>%
  arrange(chunkStrength, .by_group = TRUE) %>%
  mutate(item_ordered = factor(testItem, levels = testItem))
```

```{r}
p1 <- ggplot(dataPassive, aes(x = item_ordered)) +
  
  geom_col(aes(y = Observed, fill = "Endorsement"), width = 0.6) +
  
  geom_point(aes(y = Bayesian_predicted, shape = "Bayesian"),
             size = 3, colour = "blue",
             position = position_nudge(x = -0.15)) +
  
  geom_point(aes(y = Quantum_predicted, shape = "Quantum"),
             size = 3, colour = "red",
             position = position_nudge(x = 0.15)) +
  
  facet_wrap(~ category, scales = "free_x") +
  
  scale_fill_manual(
    name = "Data",
    values = c("Endorsement" = "grey80")
  ) +
  scale_shape_manual(
    name = "Model Predictions",
    values = c("Bayesian" = 17, "Quantum" = 15)
  ) +
  
  theme_classic(base_size = 14) +
  theme(
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  
  labs(
    x = "Item (ordered by chunk strength)",
    y = "Endorsement (probability)",
    title = "Experiment 2 (Passive)"
  )


ggsave("endorsementA_models_faceted_Passive_Exp2.png", p1, width = 12, height = 8, dpi = 300)
```

```{r}
p2 <- ggplot(dataActive, aes(x = item_ordered)) +
  
  geom_col(aes(y = Observed, fill = "Endorsement"), width = 0.6) +
  
  geom_point(aes(y = Bayesian_predicted, shape = "Bayesian"),
             size = 3, colour = "blue",
             position = position_nudge(x = -0.15)) +
  
  geom_point(aes(y = Quantum_predicted, shape = "Quantum"),
             size = 3, colour = "red",
             position = position_nudge(x = 0.15)) +
  
  facet_wrap(~ category, scales = "free_x") +
  
  scale_fill_manual(
    name = "Data",
    values = c("Endorsement" = "grey80")
  ) +
  scale_shape_manual(
    name = "Model Predictions",
    values = c("Bayesian" = 17, "Quantum" = 15)
  ) +
  
  theme_classic(base_size = 14) +
  theme(
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  
  labs(
    x = "Item (ordered by chunk strength)",
    y = "Endorsement (probability)",
    title = "Experiment 2 (Active)"
  )


ggsave("endorsementA_models_faceted_Active_Exp2.png", p2, width = 12, height = 8, dpi = 300)
```

## Conscious/Unconscious Accuracy 

```{r}
summarise(group_by(dataPs, group.fa), Mean_ConsciousAccuracy = mean(consciousAccuracy, na.rm = TRUE), StandardDeviation_ConsciousAccuracy = sd(consciousAccuracy, na.rm = TRUE))

summarise(group_by(dataPs, group.fa), Mean_UnconsciousAccuracy = mean(unconsciousAccuracy, na.rm = TRUE), StandardDeviation_UnconsciousAccuracy = sd(unconsciousAccuracy, na.rm = TRUE))
```

### Independent samples t-test

```{r}
t.test(consciousAccuracy ~ group.fa, data = dataPs)

t.test(unconsciousAccuracy ~ group.fa, data = dataPs)
```

## Accuracy of Attributed knowledge
### Accuracy of unconscious knowledge

```{r}
table(dataPs$unconsciousAccuracy)
class(dataPs$unconsciousAccuracy)
table(dataPs$unconsciousAccuracy,dataPs$group.fa)

mean_passive_unconscious <- mean(dataPs$unconsciousAccuracy[dataPs$group.fa == "Passive"], na.rm = TRUE)
mean_active_unconscious <- mean(dataPs$unconsciousAccuracy[dataPs$group.fa == "Active"], na.rm = TRUE)

overall_average <- (mean_passive_unconscious + mean_active_unconscious) / 2
overall_average
```

### Accuracy of conscious knowledge

```{r}
table(dataPs$consciousAccuracy)
class(dataPs$consciousAccuracy)
table(dataPs$consciousAccuracy,dataPs$group.fa)

mean_passive <- mean(dataPs$consciousAccuracy[dataPs$group.fa == "Passive"], na.rm = TRUE)
mean_active <- mean(dataPs$consciousAccuracy[dataPs$group.fa == "Active"], na.rm = TRUE)

overall_average <- (mean_passive + mean_active) / 2
overall_average
```

# Mixed-design ANOVA

```{r}
df_long <- pivot_longer(dataPs, cols = c(unconsciousAccuracy, consciousAccuracy),names_to = "Attribution", values_to = "Accuracy")
View(df_long)
```

### In R, the anova function expects data in a long format.

```{r}
class(df_long$Attribution)
df_long$Attribution.fa <- factor(df_long$Attribution, levels = c("unconsciousAccuracy", "consciousAccuracy"), labels = c("Implicit", "Explicit"))
class(df_long$Attribution.fa)
```

###Check for NA values in the data

```{r}
anyNA(df_long)
View(df_long)
```

###Check the number of levels in each factor

```{r}
print(paste("Number of levels in Attribution:", length(levels(df_long$Attribution.fa))))
print(paste("Number of levels in group:", length(levels(df_long$group.fa))))
```

## Performing factorial repeated measures ANOVA

```{r}
result <- aov_car(Accuracy ~ group.fa * Attribution.fa + Error(id/(Attribution.fa)), data = df_long)
summary(result)
```

# Classify participants 

```{r}
# Define thresholds considering the median of accuracy of conscious knowledge
threshold_accuracy_conscious <- median(dataPs$consciousAccuracy)  
threshold_accuracy_unconscious <- median(dataPs$unconsciousAccuracy)
threshold_accuracy_conscious
threshold_accuracy_unconscious
```

```{r}
dataPs$Group <- ifelse(dataPs$consciousAccuracy >= threshold_accuracy_conscious &  dataPs$consciousAccuracy > dataPs$unconsciousAccuracy , "Explicit knowledge", 
                       ifelse(dataPs$consciousAccuracy <= threshold_accuracy_conscious &  dataPs$consciousAccuracy < dataPs$unconsciousAccuracy, "Implicit knowledge",
                            "NA")) 
```


### Group

```{r}
table(dataPs$Group)
class(dataPs$Group)
dataPs$Group.fa <- factor(dataPs$Group)
class(dataPs$Group.fa)
table(dataPs$Group.fa)
```

```{r}
table(dataPs$Group.fa, dataPs$group.fa)
```

## Chi square 

```{r}
chisq.test(table(dataPs$Group.fa, dataPs$group.fa))
```

### Standardized Residual

```{r}
chisq <- chisq.test(table(dataPs$Group.fa, dataPs$group.fa))
chisq$stdres
```

# Plot 3

```{r}
plot_data <- dataPs %>%
  mutate(
    unconscious_pct = unconsciousAccuracy * 100,
    conscious_pct = consciousAccuracy * 100
  )

overlap_data <- plot_data %>%
  count(unconscious_pct, conscious_pct, name = "n") %>%
  filter(n > 1)

overlap_breaks <- sort(unique(overlap_data$n))

scat_1_Exp2 <- ggplot(
  plot_data,
  aes(
    x = unconscious_pct,
    y = conscious_pct,
    color = Group.fa,
    shape = group.fa
  )
) +
 
  geom_point(
    data = overlap_data,
    aes(
      x = unconscious_pct,
      y = conscious_pct,
      size = n,
      alpha = n
    ),
    inherit.aes = FALSE,
    shape = 16,
    colour = "grey35"
  ) +

  geom_point(size = 3, alpha = 0.9) +

  scale_size_area(
    max_size = 12,
    breaks = overlap_breaks,
    name = "Number at same location"
  ) +
  scale_alpha(
    range = c(0.25, 0.7),
    breaks = overlap_breaks,
    name = "Number at same location"
  ) +

  labs(
    x = "Unconscious Structural Knowledge Accuracy (%)",
    y = "Conscious Structural Knowledge Accuracy (%)",
    color = "Knowledge Attribution Group",
    shape = "Learning Condition"
  ) +
  theme_minimal()

print(scat_1_Exp2)

ggsave("scat_1_Exp2.tiff", scat_1_Exp2, width = 8, height = 5, dpi = 600, compression = "lzw")
```

# Mixed-design ANOVA
## Reshaping the data

```{r}
df_long <- pivot_longer(dataPs, cols = c(g_H, g_L, ng_H, ng_L),names_to = "RuleAdherence_ChunkStrength", values_to = "Endorsement")
```

```{r}
df_long <- separate(df_long, RuleAdherence_ChunkStrength, into = c("RuleAdherence", "ChunkStrength"), sep = "_")
```

```{r}
rule_adherence_levels <- c("g" = "Grammatical", "ng" = "Nongrammatical")
chunk_strength_levels <- c("H" = "High", "L" = "Low")
df_long$RuleAdherence <- factor(df_long$RuleAdherence, levels = names(rule_adherence_levels), labels = rule_adherence_levels)
df_long$ChunkStrength <- factor(df_long$ChunkStrength, levels = names(chunk_strength_levels), labels = chunk_strength_levels)
```

```{r}
df_long <- df_long %>% filter(Group.fa != "NA")
# Remove Unused Factor Levels
df_long$Group.fa <- droplevels(df_long$Group.fa)
table(df_long$Group.fa)
```

```{r}
anyNA(df_long)
View(df_long)
```

```{r}
print(paste("Number of levels in RuleAdherence:", length(levels(df_long$RuleAdherence))))
print(paste("Number of levels in ChunkStrength:", length(levels(df_long$ChunkStrength))))
print(paste("Number of levels in group:", length(levels(df_long$Group.fa))))
```

## Performing factorial repeated measures ANOVA

```{r}
result <- aov_car(Endorsement ~ Group.fa * RuleAdherence * ChunkStrength + Error(id/(RuleAdherence*ChunkStrength)), data = df_long)
summary(result)
```

## Pairwise t-tests 

```{r}
pairwise.t.test(df_long$Endorsement, interaction(df_long$Group.fa, df_long$RuleAdherence), p.adjust.method = "bonferroni")
pairwise.t.test(df_long$Endorsement, interaction(df_long$Group.fa,df_long$RuleAdherence, df_long$ChunkStrength), p.adjust.method = "bonferroni")
```

# Compute means and standard errors for each group

```{r}
df_long$Group.fa <- factor(
  df_long$Group.fa,
  levels = c("Implicit knowledge", "Explicit knowledge")
)

summary_data <- df_long %>%
  group_by(Group.fa, RuleAdherence, ChunkStrength) %>%
  summarise(MeanEndorsement = mean(Endorsement),
            SE = sd(Endorsement) / sqrt(n()), .groups = "drop")

print(summary_data) 
```

# Plot 4

```{r}
Violin2_exp2 <- ggplot(df_long,
       aes(x = RuleAdherence,
           y = Endorsement * 100,
           fill = ChunkStrength)) +

  geom_violin(
    alpha = 0.45,
    colour = NA,
    width = 0.75,
    position = position_dodge(width = 0.7),
    trim = FALSE
  ) +

  geom_point(
    aes(color = ChunkStrength),
    position = position_jitterdodge(
      jitter.width = 0.10,
      jitter.height = 0,
      dodge.width = 0.7
    ),
    size = 1.4,
    alpha = 0.45
  ) +

  geom_boxplot(
    width = 0.12,
    outlier.shape = NA,
    colour = "black",
    fill = "white",
    linewidth = 0.35,
    position = position_dodge(width = 0.7)
  ) +

  stat_summary(
    fun = mean,
    geom = "point",
    position = position_dodge(width = 0.7),
    size = 2.6,
    colour = "black"
  ) +

  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    width = 0.12,
    colour = "black",
    linewidth = 0.6,
    position = position_dodge(width = 0.7)
  ) +

  facet_wrap(
    ~ Group.fa,
    labeller = as_labeller(c(
      "Implicit knowledge" = "Implicit knowledge Group",
      "Explicit knowledge" = "Explicit knowledge Group"
    ))
  ) +

  labs(
    title = "(c) Experiment 2 – Knowledge Attribution",
    x = "Grammaticality Status",
    y = "Endorsement rate (%)",
    fill = "Chunk strength",
    color = "Chunk strength"
  ) +

  scale_fill_manual(values = c("#4C78A8", "#F58518")) +
  scale_color_manual(values = c("#4C78A8", "#F58518")) +

  scale_y_continuous(
    limits = c(0,100),
    breaks = seq(0,100,20),
    expand = expansion(mult = c(0,0.02))
  ) +

  theme_minimal(base_size = 14) +
  theme(
    # Text
    plot.title = element_text(size = 14, face = "bold"),
    strip.text = element_text(size = 12),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 12),

    axis.line = element_blank(),

    strip.background = element_blank(),

    panel.border = element_blank(),

    panel.grid.major.y = element_line(colour = "grey90", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),

    legend.title = element_text(size = 13),
    legend.text = element_text(size = 12),
    legend.position = "right"
  )

print(Violin2_exp2)

ggsave("Violin2_exp2.tiff", Violin2_exp2, width = 8, height = 5, dpi = 600, compression = "lzw")
```

# Individual participant performance
## 1) Passive Group
## Convert Yes/No to 1/0

```{r}
item_cols <- setdiff(names(dataIndPassive), "id")

dataIndPassive[item_cols] <- lapply(dataIndPassive[item_cols], function(x) {
  ifelse(x == "Yes", 1,
         ifelse(x == "No", 0, NA))
})
```

```{r}
table(unlist(dataIndPassive[item_cols]), useNA = "ifany")
```

## Calculate the group-level endorsement for each item

```{r}
group_endorsement <- colMeans(dataIndPassive[item_cols], na.rm = TRUE)
group_endorsement
```

## Leave-one-out individual-to-average correlation

```{r}
# Calculate leave-one-participant-out correlations
individual_correlations_loo <- numeric(nrow(dataIndPassive))

for (i in 1:nrow(dataIndPassive)) {
  
  # Participant's own responses
  participant_responses <- as.numeric(dataIndPassive[i, item_cols])
  
  # Group endorsement profile excluding participant i
  group_without_i <- colMeans(
    dataIndPassive[-i, item_cols],
    na.rm = TRUE
  )
  
  # Only use items with valid responses
  complete <- !is.na(participant_responses) &
              !is.na(group_without_i)
  
  # Correlation between participant and the remaining group
  individual_correlations_loo[i] <- cor(
    participant_responses[complete],
    group_without_i[complete],
    method = "pearson"
  )
}

# Add correlations to the dataset
dataIndPassive$individual_group_r_loo <- individual_correlations_loo
```

```{r}
summary(dataIndPassive$individual_group_r_loo)
```

# split-half analysis
## Randomly split participants into two groups

```{r}
set.seed(1234)

n_splits <- 1000

split_correlations <- numeric(n_splits)

for (s in 1:n_splits) {
  
  # Randomly assign participants to two groups
  group_assignment <- sample(
    c(rep(1, floor(nrow(dataIndPassive) / 2)),
      rep(2, ceiling(nrow(dataIndPassive) / 2)))
  )
  
  group1 <- dataIndPassive[group_assignment == 1, item_cols]
  group2 <- dataIndPassive[group_assignment == 2, item_cols]
  
  # Average endorsement for each item in each group
  endorsement1 <- colMeans(group1, na.rm = TRUE)
  endorsement2 <- colMeans(group2, na.rm = TRUE)
  
  # Correlation between the two 32-item profiles
  complete <- !is.na(endorsement1) & !is.na(endorsement2)
  
  split_correlations[s] <- cor(
    endorsement1[complete],
    endorsement2[complete],
    method = "pearson"
  )
}
```

```{r}
summary(split_correlations)
```

# Monte Carlo null simulation 

```{r}
set.seed(1234)

n_participants <- nrow(dataIndPassive)
n_sims <- 1000  # number of synthetic null datasets to generate

# "True" item-level probabilities, estimated from our actual
# full-group average (this is the null hypothesis: every
# participant shares these same probabilities, no individual
# differences)
true_p <- colMeans(dataIndPassive[, item_cols], na.rm = TRUE)
n_items <- length(true_p)

# Storage for the null distributions
null_loo_means <- numeric(n_sims)
null_split_means <- numeric(n_sims)

for (sim in 1:n_sims) {

  # Generate synthetic data: each cell is one Bernoulli
  # draw from the corresponding item's true probability. No
  # participant-level heterogeneity is built in by construction. 
  synth <- matrix(NA, nrow = n_participants, ncol = n_items)
  for (j in 1:n_items) {
    synth[, j] <- rbinom(n_participants, size = 1, prob = true_p[j])
  }
  colnames(synth) <- item_cols

  # Leave-one-out correlation 
  loo_r <- numeric(n_participants)
  for (i in 1:n_participants) {
    participant_responses <- synth[i, ]
    group_without_i <- colMeans(synth[-i, ], na.rm = TRUE)
    complete <- !is.na(participant_responses) & !is.na(group_without_i)
    loo_r[i] <- suppressWarnings(cor(
      participant_responses[complete],
      group_without_i[complete],
      method = "pearson"
    ))
  }
  null_loo_means[sim] <- mean(loo_r, na.rm = TRUE)

# Split-half correlation
  group_assignment <- sample(
    c(rep(1, floor(n_participants / 2)),
      rep(2, ceiling(n_participants / 2)))
  )
  group1 <- synth[group_assignment == 1, ]
  group2 <- synth[group_assignment == 2, ]
  endorsement1 <- colMeans(group1, na.rm = TRUE)
  endorsement2 <- colMeans(group2, na.rm = TRUE)
  complete <- !is.na(endorsement1) & !is.na(endorsement2)
  null_split_means[sim] <- suppressWarnings(cor(
    endorsement1[complete],
    endorsement2[complete],
    method = "pearson"
  ))
}

# Compare our observed statistics against these null distributions
cat("Null distribution of mean LOO correlation (no real heterogeneity):\n")
print(summary(null_loo_means))

cat("\nOur observed mean LOO correlation: ", mean(dataIndPassive$individual_group_r_loo, na.rm = TRUE), "\n")
cat("Percentile of observed value within null distribution: ",
    mean(null_loo_means <= mean(dataIndPassive$individual_group_r_loo, na.rm = TRUE)) * 100, "%\n\n")

cat("Null distribution of split-half correlation (no real heterogeneity):\n")
print(summary(null_split_means))

cat("\nOur observed mean split-half correlation: ", mean(split_correlations), "\n")
cat("Percentile of observed value within null distribution: ",
    mean(null_split_means <= mean(split_correlations)) * 100, "%\n")
```

# visual comparison

```{r}
par(mfrow = c(1, 2))
hist(null_loo_means, main = "Null: LOO correlation", xlab = "mean r",
     col = "lightgray", border = "white")
abline(v = mean(dataIndPassive$individual_group_r_loo, na.rm = TRUE), col = "red", lwd = 2)

hist(null_split_means, main = "Null: split-half correlation", xlab = "mean r",
     col = "lightgray", border = "white")
abline(v = mean(split_correlations), col = "red", lwd = 2)
```

## 2) Active Group
## Convert Yes/No to 1/0

```{r}
item_cols <- setdiff(names(dataIndActive), "id")

dataIndActive[item_cols] <- lapply(dataIndActive[item_cols], function(x) {
  ifelse(x == "Yes", 1,
         ifelse(x == "No", 0, NA))
})
```

```{r}
table(unlist(dataIndActive[item_cols]), useNA = "ifany")
```

## Calculate the group-level endorsement for each item

```{r}
group_endorsement <- colMeans(dataIndActive[item_cols], na.rm = TRUE)
group_endorsement
```

## Leave-one-out individual-to-average correlation

```{r}
# Calculate leave-one-participant-out correlations
individual_correlations_loo <- numeric(nrow(dataIndActive))

for (i in 1:nrow(dataIndActive)) {
  
  # Participant's own responses
  participant_responses <- as.numeric(dataIndActive[i, item_cols])
  
  # Group endorsement profile excluding participant i
  group_without_i <- colMeans(
    dataIndActive[-i, item_cols],
    na.rm = TRUE
  )
  
  # Only use items with valid responses
  complete <- !is.na(participant_responses) &
              !is.na(group_without_i)
  
  # Correlation between participant and the remaining group
  individual_correlations_loo[i] <- cor(
    participant_responses[complete],
    group_without_i[complete],
    method = "pearson"
  )
}

# Add correlations to the dataset
dataIndActive$individual_group_r_loo <- individual_correlations_loo
```

```{r}
summary(dataIndActive$individual_group_r_loo)
```

# split-half analysis
## Randomly split participants into two groups

```{r}
set.seed(1234)

n_splits <- 1000

split_correlations <- numeric(n_splits)

for (s in 1:n_splits) {
  
  # Randomly assign participants to two groups
  group_assignment <- sample(
    c(rep(1, floor(nrow(dataIndActive) / 2)),
      rep(2, ceiling(nrow(dataIndActive) / 2)))
  )
  
  group1 <- dataIndActive[group_assignment == 1, item_cols]
  group2 <- dataIndActive[group_assignment == 2, item_cols]
  
  # Average endorsement for each item in each group
  endorsement1 <- colMeans(group1, na.rm = TRUE)
  endorsement2 <- colMeans(group2, na.rm = TRUE)
  
  # Correlation between the two 32-item profiles
  complete <- !is.na(endorsement1) & !is.na(endorsement2)
  
  split_correlations[s] <- cor(
    endorsement1[complete],
    endorsement2[complete],
    method = "pearson"
  )
}
```

```{r}
summary(split_correlations)
```

# Monte Carlo null simulation

```{r}
set.seed(1234)

n_participants <- nrow(dataIndActive)
n_sims <- 1000  

true_p <- colMeans(dataIndActive[, item_cols], na.rm = TRUE)
n_items <- length(true_p)

null_loo_means <- numeric(n_sims)
null_split_means <- numeric(n_sims)

for (sim in 1:n_sims) {

  synth <- matrix(NA, nrow = n_participants, ncol = n_items)
  for (j in 1:n_items) {
    synth[, j] <- rbinom(n_participants, size = 1, prob = true_p[j])
  }
  colnames(synth) <- item_cols

  loo_r <- numeric(n_participants)
  for (i in 1:n_participants) {
    participant_responses <- synth[i, ]
    group_without_i <- colMeans(synth[-i, ], na.rm = TRUE)
    complete <- !is.na(participant_responses) & !is.na(group_without_i)
    loo_r[i] <- suppressWarnings(cor(
      participant_responses[complete],
      group_without_i[complete],
      method = "pearson"
    ))
  }
  null_loo_means[sim] <- mean(loo_r, na.rm = TRUE)

  group_assignment <- sample(
    c(rep(1, floor(n_participants / 2)),
      rep(2, ceiling(n_participants / 2)))
  )
  group1 <- synth[group_assignment == 1, ]
  group2 <- synth[group_assignment == 2, ]
  endorsement1 <- colMeans(group1, na.rm = TRUE)
  endorsement2 <- colMeans(group2, na.rm = TRUE)
  complete <- !is.na(endorsement1) & !is.na(endorsement2)
  null_split_means[sim] <- suppressWarnings(cor(
    endorsement1[complete],
    endorsement2[complete],
    method = "pearson"
  ))
}

cat("Null distribution of mean LOO correlation (no real heterogeneity):\n")
print(summary(null_loo_means))

cat("\nOur observed mean LOO correlation: ", mean(dataIndActive$individual_group_r_loo, na.rm = TRUE), "\n")
cat("Percentile of observed value within null distribution: ",
    mean(null_loo_means <= mean(dataIndActive$individual_group_r_loo, na.rm = TRUE)) * 100, "%\n\n")

cat("Null distribution of split-half correlation (no real heterogeneity):\n")
print(summary(null_split_means))

cat("\nOur observed mean split-half correlation: ", mean(split_correlations), "\n")
cat("Percentile of observed value within null distribution: ",
    mean(null_split_means <= mean(split_correlations)) * 100, "%\n")
```

# visual comparison

```{r}
par(mfrow = c(1, 2))
hist(null_loo_means, main = "Null: LOO correlation", xlab = "mean r",
     col = "lightgray", border = "white")
abline(v = mean(dataIndActive$individual_group_r_loo, na.rm = TRUE), col = "red", lwd = 2)

hist(null_split_means, main = "Null: split-half correlation", xlab = "mean r",
     col = "lightgray", border = "white")
abline(v = mean(split_correlations), col = "red", lwd = 2)
```






