---
title: "AGL-Exp3"
date: "`r Sys.Date()`" 
output:
  word_document:
    toc: true
  html_document:
    toc: true
    theme: cosmo
    df_print: paged
---
  
```{r setup, include = FALSE}
knitr::opts_chunk$set(comment = "", error = TRUE)
```

# Load relevant libraries 

```{r, message = FALSE}
library(readxl)
library(stargazer)
library(car)
library(reshape)
library(plyr)
library(ggplot2)
library(afex)
library(dplyr)
library(tidyr)
library(psych)
library(effectsize)
library(readr)
library(stats)
library(dplyr)
library(patchwork)
```

# Set working directory for PC

To check the current working directory path 

```{r}
getwd()
```

# Load the data
We have already removed the participants who answered the catch questions incorrectly.
```{r}
dataPs <- read_excel("AGL_Exp3.xlsx", sheet = "Behavioural")
dataStimuli <- read_excel("AGL_Exp3.xlsx", sheet = "ChunkStrength")

dataPassive <- read_excel("AGL_Exp3.xlsx", sheet = "PassiveExp3_predictions")

dataIndPassive <- read_excel("AGL_Exp3.xlsx", sheet = "IndPassive")
```

```{r}
names(dataPs)
dim(dataPs)
View(dataPs)
```

### Age

```{r}
table(dataPs$Age)
class(dataPs$Age)
max(dataPs$Age)  
min(dataPs$Age)  
mean(dataPs$Age)  
sd(dataPs$Age)   
```

### Gender

```{r}
table(dataPs$Gender)
class(dataPs$Gender)
```

### Reaction Times

```{r}
mean(dataPs$TrainingRT, na.rm = TRUE)
mean(dataPs$TestingRT, na.rm = TRUE)
```

## Runing one Sample t-test to investigate mean classification accuracy 

```{r}
t_test_allPs <- t.test(dataPs$GrammaticalityAccuracy, mu = 0.50, alternative = "greater")
t_test_allPs

g_mean <- mean(dataPs$GrammaticalityAccuracy)
g_mean
```

## Correlation Test

```{r}
cor.test(dataPs$AveCF, dataPs$AveDF, method = "pearson")
```

## Comparing overall variance of DF vs CF (Is DF a more variable construct?)

```{r}
var.test(dataPs$AveDF, dataPs$AveCF)
```

### SDs

```{r}
summaryDF <- data.frame(
  Measure = c("DF", "CF"),
  Mean = c(mean(dataPs$AveDF), mean(dataPs$AveCF)),
  SD = c(sd(dataPs$AveDF), sd(dataPs$AveCF))
)
summaryDF
```

## 1) Considering HCF and LCF 

```{r}
# Compute tail values and SD
min_val <- min(dataPs$AveCF, na.rm = TRUE)
max_val <- max(dataPs$AveCF, na.rm = TRUE)
sd_val  <- sd(dataPs$AveCF, na.rm = TRUE)

# Define tail-based cutoffs
lower_cutoff <- min_val + 1.5 * sd_val
upper_cutoff <- max_val - 1.5 * sd_val

# Classify participants
dataPs$CF_Group <- ifelse(
  dataPs$AveCF <= lower_cutoff, "LCF",
  ifelse(dataPs$AveCF >= upper_cutoff, "HCF", NA)
)

table(dataPs$CF_Group)
```

# Plot 1

```{r}
dataPs$AveCF100 <- dataPs$AveCF / 100


n_lcf <- sum(dataPs$AveCF <= lower_cutoff, na.rm = TRUE)
n_hcf <- sum(dataPs$AveCF >= upper_cutoff, na.rm = TRUE)

n_na <- sum(
  dataPs$AveCF > lower_cutoff &
    dataPs$AveCF < upper_cutoff,
  na.rm = TRUE
)

bin_width <- 0.05

low_band_start <- floor((min_val / 100) / bin_width) * bin_width
high_band_end  <- ceiling((max_val / 100) / bin_width) * bin_width

his_cf_groups_Exp3 <- ggplot(dataPs, aes(x = AveCF100)) +

  annotate(
    "rect",
    xmin = low_band_start,
    xmax = lower_cutoff / 100,
    ymin = -Inf,
    ymax = Inf,
    fill = "#4C78A8",
    alpha = 0.22
  ) +

  annotate(
    "rect",
    xmin = upper_cutoff / 100,
    xmax = high_band_end,
    ymin = -Inf,
    ymax = Inf,
    fill = "#E45756",
    alpha = 0.22
  ) +

  geom_histogram(
    binwidth = bin_width,
    boundary = 0,
    fill = "grey85",
    colour = "black",
    linewidth = 0.4
  ) +

  geom_vline(
    xintercept = lower_cutoff / 100,
    linetype = "dashed",
    linewidth = 0.9,
    colour = "#2F5D8A"
  ) +
  geom_vline(
    xintercept = upper_cutoff / 100,
    linetype = "dashed",
    linewidth = 0.9,
    colour = "#A83A3A"
  ) +

  annotate(
    "text",
    x = lower_cutoff / 100,
    y = Inf,
    label = paste0("LCF cutoff\n", round(lower_cutoff / 100, 2)),
    vjust = 1.5,
    hjust = 1.1,
    size = 3.5,
    colour = "#2F5D8A"
  ) +
  annotate(
    "text",
    x = upper_cutoff / 100,
    y = Inf,
    label = paste0("HCF cutoff\n", round(upper_cutoff / 100, 2)),
    vjust = 1.5,
    hjust = -0.1,
    size = 3.5,
    colour = "#A83A3A"
  ) +

  annotate(
    "text",
    x = (low_band_start + lower_cutoff / 100) / 2,
    y = Inf,
    label = paste0("LCF\nn = ", n_lcf),
    vjust = 4.5,
    size = 3.8,
    fontface = "bold",
    colour = "#2F5D8A"
  ) +

  annotate(
    "text",
    x = (upper_cutoff / 100 + high_band_end) / 2,
    y = Inf,
    label = paste0("HCF\nn = ", n_hcf),
    vjust = 4.5,
    size = 3.8,
    fontface = "bold",
    colour = "#A83A3A"
  ) +

  annotate(
    "text",
    x = (lower_cutoff / 100 + upper_cutoff / 100) / 2,
    y = Inf,
    label = paste0("Not assigned\nn = ", n_na),
    vjust = 4.5,
    size = 3.5,
    colour = "grey25"
  ) +

  scale_x_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, 0.1),
    labels = label_number(accuracy = 0.1)
  ) +
  labs(
    title = " (a) Distribution of Average Conjunction-Fallacy Scores",
    x = "Average conjunction-fallacy susceptibility",
    y = "Number of participants"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 11, face = "bold"),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 11),
    plot.caption = element_text(hjust = 0, size = 9),
    legend.position = "none"
  )

his_cf_groups_Exp3

ggsave(
  "his_cf_groups_Exp3.tiff",
  his_cf_groups_Exp3,
  width = 6,
  height = 4,
  dpi = 600,
  compression = "lzw"
)
```

### CF Group

```{r}
class(dataPs$CF_Group)
dataPs$CF_Group.fa <- factor(dataPs$CF_Group)
table(dataPs$CF_Group.fa)
class(dataPs$CF_Group.fa)
```

```{r}
mean_rts <- dataPs %>%
  group_by(CF_Group) %>%
  summarise(
    Mean_TrainingRT = mean(TrainingRT, na.rm = TRUE),
    Mean_TestingRT  = mean(TestingRT, na.rm = TRUE),
    Mean_CF         = mean(AveCF, na.rm = TRUE),
    Mean_DF         = mean(AveDF, na.rm = TRUE),
    Count           = n()
  )

mean_rts
```

## t‑test

```{r}
t.test(AveCF ~ CF_Group.fa, data = dataPs, var.equal = FALSE)
```

# ANOVA
## Reshaping the data

```{r}
dataPs <- rename(
  dataPs,
  G_H  = GH,
  G_L  = GL,
  NG_H = NGH,
  NG_L = NGL
)
names(dataPs)
```

```{r}
df_long <- pivot_longer(dataPs, cols = c(G_H, G_L, NG_H, NG_L),names_to = "Grammaticality_ChunkStrength", values_to = "Endorsement")
```

```{r}
df_long <- separate(df_long, Grammaticality_ChunkStrength, into = c("Grammaticality", "ChunkStrength"), sep = "_")
```

```{r}
Grammaticality_levels <- c("G" = "Grammatical", "NG" = "Nongrammatical")
chunk_strength_levels <- c("H" = "High", "L" = "Low")
df_long$Grammaticality <- factor(df_long$Grammaticality, levels = names(Grammaticality_levels), labels = Grammaticality_levels)
df_long$ChunkStrength <- factor(df_long$ChunkStrength, levels = names(chunk_strength_levels), labels = chunk_strength_levels)
```

```{r}
str(df_long$Grammaticality)
str(df_long$ChunkStrength)
```

### Repeated-measures 2x2 ANOVA

```{r}
result_within <- aov_car(Endorsement ~ Grammaticality * ChunkStrength + Error(Ps/(Grammaticality * ChunkStrength)), data = df_long)
summary(result_within)
```

## Pairwise t-tests 

```{r}
pairwise.t.test(df_long$Endorsement, interaction(df_long$ChunkStrength, df_long$Grammaticality), p.adjust.method = "bonferroni")
```

# Compute means and standard deviations

```{r}
summary_data <- df_long %>%
  group_by(Grammaticality, ChunkStrength) %>%
  summarise(
    MeanEndorsement = mean(Endorsement),
    SD = sd(Endorsement),
    N = n(),
    SE = SD / sqrt(N),
    .groups = "drop"
  )

print(summary_data)
```

# Plot 2

```{r}
Violin_exp3 <- ggplot(df_long,
       aes(x = Grammaticality,
           y = Endorsement * 100,
           fill = ChunkStrength)) +

  geom_violin(
    alpha = 0.45,
    colour = NA,
    width = 0.75,
    position = position_dodge(width = 0.7),
    trim = FALSE
  ) +

  geom_point(
    aes(color = ChunkStrength),
    position = position_jitterdodge(
      jitter.width = 0.10,
      jitter.height = 0,
      dodge.width = 0.7
    ),
    size = 1.4,
    alpha = 0.45
  ) +

  geom_boxplot(
    width = 0.12,
    outlier.shape = NA,
    colour = "black",
    fill = "white",
    linewidth = 0.35,
    position = position_dodge(width = 0.7)
  ) +

  stat_summary(
    fun = mean,
    geom = "point",
    position = position_dodge(width = 0.7),
    size = 2.6,
    colour = "black"
  ) +

  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    width = 0.12,
    colour = "black",
    linewidth = 0.6,
    position = position_dodge(width = 0.7)
  ) +

  labs(
    title = "(d) Experiment 3",
    x = "Grammaticality Status",
    y = "Endorsement rate (%)",
    fill = "Chunk strength",
    color = "Chunk strength"
  ) +

  scale_fill_manual(values = c("#4C78A8", "#F58518")) +
  scale_color_manual(values = c("#4C78A8", "#F58518")) +

  scale_y_continuous(
    limits = c(0,100),
    breaks = seq(0,100,20),
    expand = expansion(mult = c(0,0.02))
  ) +

  theme_minimal(base_size = 14) +
  theme(

    plot.title = element_text(size = 14, face = "bold"),
    strip.text = element_text(size = 12),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 12),

    axis.line = element_blank(),

    strip.background = element_blank(),

    panel.border = element_blank(),

    panel.grid.major.y = element_line(colour = "grey90", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),

    legend.title = element_text(size = 13),
    legend.text = element_text(size = 12),
    legend.position = "right"
  )

print(Violin_exp3)

ggsave("Violin_exp3.tiff", Violin_exp3, width = 5, height = 5, dpi = 600, compression = "lzw")
```

### Calculate Mean RTs by Group

```{r}
mean_rts <- df_long %>%
  group_by(CF_Group) %>%
  summarise(
    Mean_TrainingRT = mean(TrainingRT, na.rm = TRUE),
    Mean_TestingRT  = mean(TestingRT, na.rm = TRUE),
    Count           = n()
  )
```

# Mixed-design ANOVA
## Reshaping the data

```{r}
#Remove rows where CF_Group.fa is "NA" 
df_long <- df_long %>% filter(CF_Group.fa != "NA")

df_long$Group.fa <- droplevels(df_long$CF_Group.fa)
table(df_long$Group.fa)
```

```{r}
anyNA(df_long)
View(df_long)
```

```{r}
print(paste("Number of levels in Grammaticality:", length(levels(df_long$Grammaticality))))
print(paste("Number of levels in ChunkStrength:", length(levels(df_long$ChunkStrength))))
print(paste("Number of levels in group:", length(levels(df_long$CF_Group.fa))))
```

## Performing factorial repeated measures ANOVA

```{r}
result <- aov_car(Endorsement ~ Group.fa * Grammaticality * ChunkStrength + Error(Ps/(Grammaticality*ChunkStrength)), data = df_long)
summary(result)
```

## 2) Considering HDF and LDF 

```{r}
# Compute tail values and SD
min_val <- min(dataPs$AveDF, na.rm = TRUE)
max_val <- max(dataPs$AveDF, na.rm = TRUE)
sd_val  <- sd(dataPs$AveDF, na.rm = TRUE)

# Define tail-based cutoffs
lower_cutoff <- min_val + 1.5 * sd_val
upper_cutoff <- max_val - 1.5 * sd_val

# Classify participants
dataPs$DF_Group <- ifelse(
  dataPs$AveDF <= lower_cutoff, "LDF",
  ifelse(dataPs$AveDF >= upper_cutoff, "HDF", NA)
)

# Display frequencies
table(dataPs$DF_Group)
```

# Plot 3

```{r}
dataPs$AveDF100 <- dataPs$AveDF / 100
dataPs$AveDF100 <- floor(dataPs$AveDF100 / 0.05) * 0.05

n_ldf <- sum(dataPs$AveDF <= lower_cutoff, na.rm = TRUE)
n_hdf <- sum(dataPs$AveDF >= upper_cutoff, na.rm = TRUE)

n_na <- sum(
  dataPs$AveDF > lower_cutoff &
    dataPs$AveDF < upper_cutoff,
  na.rm = TRUE
)

bin_width <- 0.05

low_band_start <- floor((min_val / 100) / bin_width) * bin_width
high_band_end  <- ceiling((max_val / 100) / bin_width) * bin_width

his_df_groups_Exp3 <- ggplot(dataPs, aes(x = AveDF100)) +

  annotate(
    "rect",
    xmin = low_band_start,
    xmax = lower_cutoff / 100,
    ymin = -Inf,
    ymax = Inf,
    fill = "#4C78A8",
    alpha = 0.22
  ) +

  annotate(
    "rect",
    xmin = upper_cutoff / 100,
    xmax = high_band_end,
    ymin = -Inf,
    ymax = Inf,
    fill = "#E45756",
    alpha = 0.22
  ) +

  geom_histogram(
    binwidth = bin_width,
    boundary = 0,
    fill = "grey85",
    colour = "black",
    linewidth = 0.4
  ) +

  geom_vline(
    xintercept = lower_cutoff / 100,
    linetype = "dashed",
    linewidth = 0.9,
    colour = "#2F5D8A"
  ) +
  geom_vline(
    xintercept = upper_cutoff / 100,
    linetype = "dashed",
    linewidth = 0.9,
    colour = "#A83A3A"
  ) +

  annotate(
    "text",
    x = lower_cutoff / 100,
    y = Inf,
    label = paste0("LDF cutoff\n", round(lower_cutoff / 100, 2)),
    vjust = 1.5,
    hjust = 1.1,
    size = 3.5,
    colour = "#2F5D8A"
  ) +
  annotate(
    "text",
    x = upper_cutoff / 100,
    y = Inf,
    label = paste0("HDF cutoff\n", round(upper_cutoff / 100, 2)),
    vjust = 1.5,
    hjust = -0.1,
    size = 3.5,
    colour = "#A83A3A"
  ) +

  annotate(
    "text",
    x = (low_band_start + lower_cutoff / 100) / 2,
    y = Inf,
    label = paste0("LDF\nn = ", n_ldf),
    vjust = 4.5,
    size = 3.8,
    fontface = "bold",
    colour = "#2F5D8A"
  ) +

  annotate(
    "text",
    x = (upper_cutoff / 100 + high_band_end) / 2,
    y = Inf,
    label = paste0("HDF\nn = ", n_hdf),
    vjust = 4.5,
    size = 3.8,
    fontface = "bold",
    colour = "#A83A3A"
  ) +

  annotate(
    "text",
    x = (lower_cutoff / 100 + upper_cutoff / 100) / 2,
    y = Inf,
    label = paste0("Not assigned\nn = ", n_na),
    vjust = 4.5,
    size = 3.5,
    colour = "grey25"
  ) +

  scale_x_continuous(
    limits = c(0, 1.25),
    breaks = seq(0, 1.3, 0.1),
    labels = label_number(accuracy = 0.1)
  ) +
  labs(
    title = "(b) Distribution of Average Disjunction-Fallacy Scores",
    x = "Average disjunction-fallacy susceptibility",
    y = "Number of participants"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 11, face = "bold"),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 11),
    plot.caption = element_text(hjust = 0, size = 9),
    legend.position = "none"
  )

his_df_groups_Exp3

ggsave(
  "his_df_groups_Exp3.tiff",
  his_df_groups_Exp3,
  width = 6,
  height = 4,
  dpi = 600,
  compression = "lzw"
)
```

### DF Group

```{r}
table(dataPs$DF_Group)
class(dataPs$DF_Group)
dataPs$DF_Group.fa <- factor(dataPs$DF_Group)
table(dataPs$DF_Group.fa)
class(dataPs$DF_Group.fa)
```

### Calculate Mean RTs by Group

```{r}
mean_rts <- dataPs %>%
  group_by(DF_Group) %>%
  summarise(
    Mean_TrainingRT = mean(TrainingRT, na.rm = TRUE),
    Mean_TestingRT  = mean(TestingRT, na.rm = TRUE),
    Count           = n()
  )

mean_rts
```

## t‑test

```{r}
t.test(AveDF ~ DF_Group.fa, data = dataPs, var.equal = FALSE)
```

# Mixed-design ANOVA
## Reshaping the data

```{r}
df_long <- pivot_longer(dataPs, cols = c(G_H, G_L, NG_H, NG_L),names_to = "Grammaticality_ChunkStrength", values_to = "Endorsement")
```

```{r}
df_long <- separate(df_long, Grammaticality_ChunkStrength, into = c("Grammaticality", "ChunkStrength"), sep = "_")
```

```{r}
Grammaticality_levels <- c("G" = "Grammatical", "NG" = "Nongrammatical")
chunk_strength_levels <- c("H" = "High", "L" = "Low")
df_long$Grammaticality <- factor(df_long$Grammaticality, levels = names(Grammaticality_levels), labels = Grammaticality_levels)
df_long$ChunkStrength <- factor(df_long$ChunkStrength, levels = names(chunk_strength_levels), labels = chunk_strength_levels)
```

```{r}
str(df_long$Grammaticality)
str(df_long$ChunkStrength)
```

```{r}
## Remove rows where CF_Group.fa is "NA" 
df_long <- df_long %>% filter(DF_Group.fa != "NA")

df_long$DF_Group.fa <- droplevels(df_long$DF_Group.fa)
table(df_long$DF_Group.fa)
```

```{r}
anyNA(df_long)
View(df_long)
```

```{r}
print(paste("Number of levels in Grammaticality:", length(levels(df_long$Grammaticality))))
print(paste("Number of levels in ChunkStrength:", length(levels(df_long$ChunkStrength))))
print(paste("Number of levels in group:", length(levels(df_long$DF_Group.fa))))
```

## Performing factorial repeated measures ANOVA

```{r}
result <- aov_car(Endorsement ~ DF_Group.fa * Grammaticality * ChunkStrength + Error(Ps/(Grammaticality*ChunkStrength)), data = df_long)
summary(result)
```

```{r}
write.csv(dataPs, "dataPs_Exp3.csv", row.names = FALSE)
```

# RSS
## Min–max normalise: chunkStrength

```{r}
# Passive data
dataPassive$chunkStrength_norm <- (
  dataPassive$chunkStrength - min(dataPassive$chunkStrength)
) / (max(dataPassive$chunkStrength) - min(dataPassive$chunkStrength))

```

## Define an RSS function

```{r}
rss <- function(observed, predicted) {
  sum((observed - predicted)^2)
}
```

### Compute RSS for passive learning 

```{r}
RSS_Bayesian <- rss(dataPassive$Observed, dataPassive$Bayesian_predicted)
RSS_Quantum <- rss(dataPassive$Observed, dataPassive$Quantum_predicted)
RSS_chunkStrength <- rss(dataPassive$Observed, dataPassive$chunkStrength_norm)
RSS_grammaticality <- rss(dataPassive$Observed, dataPassive$grammaticality)
```

```{r}
RSS_results_Passive <- data.frame(
  Predictor = c("Bayesian", "Quantum", "Chunk Strength (norm)", "Grammaticality"),
  RSS = c(RSS_Bayesian, RSS_Quantum, RSS_chunkStrength, RSS_grammaticality)
)

RSS_results_Passive
```

# Plot 4

### Create the category variable

```{r}
# Passive
dataPassive$category <- cut(
  seq_len(nrow(dataPassive)),
  breaks = c(0, 8, 16, 24, 32),
  labels = c("GH", "GL", "NGH", "NGL")
)

```

```{r}
# Passive
dataPassive <- dataPassive %>%
  group_by(category) %>%
  arrange(chunkStrength, .by_group = TRUE) %>%
  mutate(item_ordered = factor(testItem, levels = testItem))

```

```{r}
p1 <- ggplot(dataPassive, aes(x = item_ordered)) +
  
  geom_col(aes(y = Observed, fill = "Endorsement"), width = 0.6) +
  
  geom_point(aes(y = Bayesian_predicted, shape = "Bayesian"),
             size = 3, colour = "blue",
             position = position_nudge(x = -0.15)) +
  
  geom_point(aes(y = Quantum_predicted, shape = "Quantum"),
             size = 3, colour = "red",
             position = position_nudge(x = 0.15)) +
  
  facet_wrap(~ category, scales = "free_x") +
  
  scale_fill_manual(
    name = "Data",
    values = c("Endorsement" = "grey80")
  ) +
  scale_shape_manual(
    name = "Model Predictions",
    values = c("Bayesian" = 17, "Quantum" = 15)
  ) +
  
  theme_classic(base_size = 14) +
  theme(
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  
  labs(
    x = "Item (ordered by chunk strength)",
    y = "Endorsement (probability)",
    title = "Experiment 3 (Passive)"
  )

ggsave("endorsementA_models_faceted_Passive_Exp3.png", p1, width = 12, height = 8, dpi = 300)
```


# Individual participant performance
## Convert Yes/No to 1/0

```{r}
item_cols <- setdiff(names(dataIndPassive), "id")

dataIndPassive[item_cols] <- lapply(dataIndPassive[item_cols], function(x) {
  ifelse(x == "Yes", 1,
         ifelse(x == "No", 0, NA))
})
```

```{r}
table(unlist(dataIndPassive[item_cols]), useNA = "ifany")
```

## Calculate the group-level endorsement for each item

```{r}
group_endorsement <- colMeans(dataIndPassive[item_cols], na.rm = TRUE)
group_endorsement
```

## Leave-one-out individual-to-average correlation

```{r}
# Calculate leave-one-participant-out correlations
individual_correlations_loo <- numeric(nrow(dataIndPassive))

for (i in 1:nrow(dataIndPassive)) {
  
  # Participant's own responses
  participant_responses <- as.numeric(dataIndPassive[i, item_cols])
  
  # Group endorsement profile excluding participant i
  group_without_i <- colMeans(
    dataIndPassive[-i, item_cols],
    na.rm = TRUE
  )
  
  # Only use items with valid responses
  complete <- !is.na(participant_responses) &
              !is.na(group_without_i)
  
  # Correlation between participant and the remaining group
  individual_correlations_loo[i] <- cor(
    participant_responses[complete],
    group_without_i[complete],
    method = "pearson"
  )
}

# Add correlations to the dataset
dataIndPassive$individual_group_r_loo <- individual_correlations_loo
```

```{r}
summary(dataIndPassive$individual_group_r_loo)
```

# split-half analysis
## Randomly split participants into two groups

```{r}
set.seed(1234)

n_splits <- 1000

split_correlations <- numeric(n_splits)

for (s in 1:n_splits) {
  
  # Randomly assign participants to two groups
  group_assignment <- sample(
    c(rep(1, floor(nrow(dataIndPassive) / 2)),
      rep(2, ceiling(nrow(dataIndPassive) / 2)))
  )
  
  group1 <- dataIndPassive[group_assignment == 1, item_cols]
  group2 <- dataIndPassive[group_assignment == 2, item_cols]
  
  # Average endorsement for each item in each group
  endorsement1 <- colMeans(group1, na.rm = TRUE)
  endorsement2 <- colMeans(group2, na.rm = TRUE)
  
  # Correlation between the two 32-item profiles
  complete <- !is.na(endorsement1) & !is.na(endorsement2)
  
  split_correlations[s] <- cor(
    endorsement1[complete],
    endorsement2[complete],
    method = "pearson"
  )
}
```

```{r}
summary(split_correlations)
```

# Monte Carlo null simulation

```{r}
set.seed(1234)

n_participants <- nrow(dataIndPassive)
n_sims <- 1000  # number of synthetic null datasets to generate

# "True" item-level probabilities, estimated from our actual
# full-group average (this is the null hypothesis: every
# participant shares these same probabilities, no individual
# differences)
true_p <- colMeans(dataIndPassive[, item_cols], na.rm = TRUE)
n_items <- length(true_p)

# Storage for the null distributions
null_loo_means <- numeric(n_sims)
null_split_means <- numeric(n_sims)

for (sim in 1:n_sims) {

  # Generate synthetic data: each cell is one Bernoulli
  # draw from the corresponding item's true probability. No
  # participant-level heterogeneity is built in by construction. 
  synth <- matrix(NA, nrow = n_participants, ncol = n_items)
  for (j in 1:n_items) {
    synth[, j] <- rbinom(n_participants, size = 1, prob = true_p[j])
  }
  colnames(synth) <- item_cols

  # Leave-one-out correlation 
  loo_r <- numeric(n_participants)
  for (i in 1:n_participants) {
    participant_responses <- synth[i, ]
    group_without_i <- colMeans(synth[-i, ], na.rm = TRUE)
    complete <- !is.na(participant_responses) & !is.na(group_without_i)
    loo_r[i] <- suppressWarnings(cor(
      participant_responses[complete],
      group_without_i[complete],
      method = "pearson"
    ))
  }
  null_loo_means[sim] <- mean(loo_r, na.rm = TRUE)

# Split-half correlation
  group_assignment <- sample(
    c(rep(1, floor(n_participants / 2)),
      rep(2, ceiling(n_participants / 2)))
  )
  group1 <- synth[group_assignment == 1, ]
  group2 <- synth[group_assignment == 2, ]
  endorsement1 <- colMeans(group1, na.rm = TRUE)
  endorsement2 <- colMeans(group2, na.rm = TRUE)
  complete <- !is.na(endorsement1) & !is.na(endorsement2)
  null_split_means[sim] <- suppressWarnings(cor(
    endorsement1[complete],
    endorsement2[complete],
    method = "pearson"
  ))
}

# Compare our observed statistics against these null distributions
cat("Null distribution of mean LOO correlation (no real heterogeneity):\n")
print(summary(null_loo_means))

cat("\nOur observed mean LOO correlation: ", mean(dataIndPassive$individual_group_r_loo, na.rm = TRUE), "\n")
cat("Percentile of observed value within null distribution: ",
    mean(null_loo_means <= mean(dataIndPassive$individual_group_r_loo, na.rm = TRUE)) * 100, "%\n\n")

cat("Null distribution of split-half correlation (no real heterogeneity):\n")
print(summary(null_split_means))

cat("\nOur observed mean split-half correlation: ", mean(split_correlations), "\n")
cat("Percentile of observed value within null distribution: ",
    mean(null_split_means <= mean(split_correlations)) * 100, "%\n")
```

# visual comparison

```{r}
par(mfrow = c(1, 2))
hist(null_loo_means, main = "Null: LOO correlation", xlab = "mean r",
     col = "lightgray", border = "white")
abline(v = mean(dataIndPassive$individual_group_r_loo, na.rm = TRUE), col = "red", lwd = 2)

hist(null_split_means, main = "Null: split-half correlation", xlab = "mean r",
     col = "lightgray", border = "white")
abline(v = mean(split_correlations), col = "red", lwd = 2)
```




